---
output:
  pdf_document: default
  html_document: default
---
# Covid-19 Death Rates
## Determining the variables that increase chances of death
### By: Olusegun Stephens


## Retrieve Data

We download and view the John's Hopkins data of COVID-19 cases from January to February of 2020. 
```{r}
rm(list=ls()) #remove remaining variables
library(Hmisc) #import

data <- read.csv("~/Coding/data analysis/Projects/Covid Data R/COVID19_line_list_data.csv")
describe(data)
```
## Cleaning Data

Looking at the data, we see some discrepancies on how the data was recorded, specifically in the death column. Since we know that 0 means the person survived, we can mark dates of death as 1 under another column:

```{r}
data$death_dummy <- as.integer(data$death !=0)
unique(data$death_dummy)
```
## Analyzing Data

Now we can confidently find our death rate.

```{r}
sum(data$death_dummy) / nrow(data)
```
About 5.8% of COVID-19 cases result in death. Are there any variables that could increase the chance of death?

We can check age to see if the older population are more at risk. We first need to find the average age for the living and dead cases:

```{r}
dead = subset(data, death_dummy == 1)
alive = subset(data, death_dummy == 0)
mean(dead$age, na.rm = TRUE)
mean(alive$age, na.rm = TRUE)
```
The average age of death cases is 68.6, while living cases are on average 48.1. There is a 20 year difference in the ages, but let's check to see if this is statistically significant.

```{r}
t.test(alive$age, dead$age, alternative = "two.sided", conf.level =  0.99)
```
We have a p-value of 2.2e-16, or less than 0.05, making these findings significant. 

We can check to see if gender plays a role in the death risk as well:

```{r}
men = subset(data, gender == 'male')
women = subset(data, gender == 'female')
mean(men$death_dummy, na.rm = TRUE) 
mean(women$death_dummy, na.rm = TRUE)

t.test(men$death_dummy, women$death_dummy, alternative = "two.sided", conf.level =  0.99)
```
Men make up 8.5% of deaths and women 3.7%. We have a p-value of 0.002, which again is less than 0.05. Gender seems to play a role in the chances of surviving COVID-19.

## Conclusion

With our findings, it could be suggested that older men have a higher risk of death when they contract COVID-19. Though, overall the illness has a low death rate at 5%.
